---
title: "Statistical Programming with R"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

# Intro {.sidebar}

This dashboard covers the materials for the third course in R programming. Course held in Sarajevo, January 27-20, 2020. The course websites for the previous courses can be found [here](http://peterstoltze.dk/RfOS/) and [here](https://joliencremers.github.io/Course-Materials-Bosnia/).

---

Instructors: 

 - Aksel Thomsen (akt@dst.dk)
 - Signe Hermann (she@dst.dk)
 
 
---

Part of the material for this course is adapted from [Gerko Vink](https://github.com/gerkovink/R)


---
# Quick Overview

## Column 1

### Outline

This follow-up course deals with the following topics:

1. Sampling
1. Estimation and calibration
1. Functional programming
1. Preparing large data for analysis
1. Analysis of large data

Previous experience with R is required.

We adapt the course as we go. To ensure that you work with the latest iteration of the course materials, we advice all course participants to access the materials online.

If you want access to the underlying R files, that creates the presentation and more, they can be accesed from [GitHub](https://github.com/akselthomsen/IR-Bosnia-III).

*Update 29 January*: Due to time constraints, the two last subjects has been combined into a smaller section on analysing large data.

## Column 2

### Course schedule 

| Time        | Topic                                           |
|:------------|:------------------------------------------------|
| *Tuesday*   |                                                 |
| 09.00-10.30 | Intro & Repetition (A)                          |
| 10.30-12.00 | Sampling (B)                                    |
|             | Break                                           |    
| 13.00-15.30 | Sampling *continued*                            |
|             | Estimation and calibration (C)                  |
| *Wednesday* |                                                 |
| 09.00-12.00 | Estimation and calibration *continued*          |
|             | Functional Programming (D)                      |
|             | *Break*                                         |             
| 13.00-15.30 | Functional Programming *continued*              |
| *Thursday*  |                                                 |
| 09.00-11.00 | Analysing large data (F)                        |
| 11.00-12.00 | Finish mission report                           |

# How to prepare

## Column 1

### Welcome

Welcome to the third course.
In general the requirements are the same, as for the two previous.
We might add some extra packages a long the way, but otherwise if everything worked last time - it will also work this time.

On thursday we wil be working with some large data sets than usual in these courses. We will provide links to them on the course webpage, but we will also have them available on several USB sticks to save you the trouble of downloading several GBs.

/Aksel og Signe

### **0. System requirements**

Bring a laptop computer to the course and make sure that you have full write access and administrator rights to the machine. We will explore programming and compiling in this course. This means that you need full access to your machine. Some corporate laptops come with limited access for their users, we therefore advice you to bring a personal laptop computer, if you have one. 

### **1. Install `R`**
`R` can be obtained [here](https://cran.r-project.org). We won't use `R` directly in the course, but rather call `R` through `RStudio`. Therefore it needs to be installed. 

### **2. Install `RStudio` Desktop**

Rstudio is an Integrated Development Environment (IDE). It can be obtained as stand-alone software [here](https://www.rstudio.com/products/rstudio/download/#download). The free and open source `RStudio Desktop` version is sufficient.

### **3. Start RStudio and install the following packages. **

Execute the following lines of code in the console window:

```{r eval=FALSE, echo = TRUE}
install.packages(c("tidyverse", "micemd", "jomo", "pan", 
                 "lme4", "knitr", "rmarkdown", "plotly",
                 "devtools", "boot", "class", "car", "MASS", 
                 "ISLR", "DAAG", "mice", "mitml", "miceadds",
                 "Ecdat", "Ecfun", "MEMSS", "VIM", "simputation",
                 "naniar","visdat", "UpSetR", "DAAG", "magrittr",
                 "haven", "Matrix", "lattice", "data.table", "grid",
                 "colorspace", "stringi", "stringdist", "editrules", 
                 "deducorrect", "rex"), 
                 dependencies = TRUE)
```

Just copy and paste the installation command and press the return key. When asked 

```{r eval = FALSE, echo = TRUE}
Do you want to install from sources the package which needs 
compilation? (Yes/no/cancel)
```

type `Yes` in the console and press the return key. 

# Tuesday

## Column 1

### Tuesday's materials

- Part A: Repetition

    - [Lecture A](Contents/Material/Part A/Lecture_A.html){target="_blank"}
    - Data:
       - [mtcars.rds](Contents/Material/Part A/mtcars.rds){target="_blank"}
       - [mtcars5.csv](Contents/Material/Part A/mtcars5.csv){target="_blank"}
       - [mtcars5.xlsx](Contents/Material/Part A/mtcars5.xlsx){target="_blank"}
    
- Part B: Sampling

    - [Lecture B](Contents/Material/Part B/Sampling.html){target="_blank"}
    - [Exercise B](Contents/Material/Part B/Sampling_Exercises.html){target="_blank"}
    - [Solution B](Contents/Material/Part B/Sampling_Exercises_Solutions.html){target="_blank"}
    - Data 
       - Lecture:
          - [mydata_example.Rda](Contents/Material/Part B/mydata_example.Rda){target="_blank"}
          - [mysample_opt.Rda](Contents/Material/Part B/mysample_opt.Rda){target="_blank"}
          - [mysample_prop.Rda](Contents/Material/Part B/mysample_prop.Rda){target="_blank"}
       - Exercise:
          - [mydata_estimation.Rda](Contents/Material/Part B/mydata_estimation.Rda){target="_blank"}
       - Solution:
          - [mysampopt.Rda](Contents/Material/Part B/mysampopt.Rda){target="_blank"}
          - [mysampprop.Rda](Contents/Material/Part B/mysampprop.Rda){target="_blank"}
          - [mysampweigh.Rda](Contents/Material/Part B/mysampweigh.Rda){target="_blank"}
    
- Part C: Estimation and calibration

    - [Lecture C](Contents/Material/Part C/Estimation.html){target="_blank"}
    - [Exercise C](Contents/Material/Part C/Estimation_Exercises.html){target="_blank"}
    - [Solution C](Contents/Material/Part C/Estimation_Exercises.html){target="_blank"}
    - Data: Reuse from Part B

## Column 2

### Useful references
- [tidyverse.com](https://www.tidyverse.org/): Everything about the tidyverse and their packages
- [RStudio Cheat Sheets](https://rstudio.com/resources/cheatsheets/): Collection of very good cheat sheets
- [Tidy data](https://www.jstatsoft.org/article/view/v059i10): Article explaining the tidy data concept and best practises for structuring data in R. Written by Hadley Wickham, Chief Scientist at Rstudio.

# Wednesday

## Column 1

### Wednesday's materials

- Part C: Estimation and calibration (continued from yesterday)

    - [Lecture C](Contents/Material/Part C/Estimation.html){target="_blank"}
    - [Exercise C](Contents/Material/Part C/Estimation_Exercises.html){target="_blank"}
    - [Solution C](Contents/Material/Part C/Estimation_Exercises_Solutions.html){target="_blank"}
    - Data: Reuse from Part B
    
- Part D: Functional Programming

    - [Lecture D](Contents/Material/Part D/Lecture_D.html){target="_blank"}
    - [Exercise D1](Contents/Material/Part D/Exercise_D1.html){target="_blank"}
    - [Solution D1](Contents/Material/Part D/Solutions_D1.html){target="_blank"}
    - [Exercise D2](Contents/Material/Part D/Exercise_D2.html){target="_blank"}
    - [Solution D2](Contents/Material/Part D/Solution_D2.html){target="_blank"}
    - Data:
       - [cars_1.csv](Contents/Material/Part D/cars_1.csv){target="_blank"}
       - [cars_2.csv](Contents/Material/Part D/cars_2.csv){target="_blank"}
       - [cars_3.csv](Contents/Material/Part D/cars_3.csv){target="_blank"}
       - [cars_4.csv](Contents/Material/Part D/cars_4.csv){target="_blank"}
       - [cars_5.csv](Contents/Material/Part D/cars_5.csv){target="_blank"}
       - [mydata_example.Rda](Contents/Material/Part D/mydata_example.Rda){target="_blank"}

## Column 2

### Useful references
- [survey package](https://CRAN.R-project.org/package=survey): Link to the CRAN documentation
- [A survey analysis example](https://cran.r-project.org/web/packages/survey/vignettes/survey.pdf) by Thomas Lumley: Introduction vignette to the `survey` package
- [Basic R examples](https://www.coursera.org/lecture/data-collection-analytics-project/basic-r-examples-eYiuT): Coursera course on using the `survey` package
- [Cousera](https://www.coursera.org/): General link

# Thursday

## Column 1

### Thursdays's materials

- Part E: Analysing large data

    - [Lecture E](Contents/Material/Part E/Lecture.html){target="_blank"}
    - [Exercise E](Contents/Material/Part E/Exercise.html){target="_blank"}
    - [Solution E](Contents/Material/Part E/Exercise.html){target="_blank"}

## Column 2

### Useful references
- [add]()

# Further studies

## Column 1

### What to do after the course

The following references are currently available for free, either as pdfs or as extensive webpages (written with [RMarkdown](https://rmarkdown.rstudio.com/) and [bookdown](https://bookdown.org/)). They are all very useful and we highly recommend them. 

- [R for Data Science](https://r4ds.had.co.nz): written by Hadley Wickham and Garrett Grolemund this book relies almost exclusively on the [tidyverse](https://www.tidyverse.org/) approach to data analysis. Many highly effective tools will be right at your fingertips after reading this book and working through the many exercises.
- [Hands-On Programming with R](https://rstudio-education.github.io/hopr/): a great read by Garrett Grolemund emphasizing programming techniques with R. 
- [Advanced R](https://adv-r.hadley.nz/): You want to gain deeper knowledge of R and you want to learn from one of the most influential R contributors. This one is for you!
- [Introduction to Statistical Learning](http://faculty.marshall.usc.edu/gareth-james/ISL/): an introductory book on statistical learning, with applications in R. The R code is somewhat old-style and you might be able to find newer packages for the tasks, but ths is a solid read well worth the effort.
- [Data Analysis and Graphics Using R](http://www.pindex.com/uploads/post_docs/Maindonald%20Data%20Analysis%20and%20Graphics(PINDEX-DOC-6953).pdf): a detailed book that covers a lot about categorical data analysis and fitting `glm`s in `R`.
- [Happy Git and GitHub for the useR ](https://happygitwithr.com/index.html): a great introduction to version control using Git and GitHub together with RStudio. Written by Jenny Bryan in a very concise style. Highly recommended!
- [Tidyverse Cheat Sheets](https://rstudio.com/resources/cheatsheets/): a collection of excellent cheat sheets when using tidyverse functions. Highly recomendable!
- [Combining and Analyzing Complex Data](https://www.coursera.org/learn/data-collection-analytics-project): Check out this Cousera course

<!-- ## Column 2 -->

