---
title: "Functional Programming"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output:
  ioslides_presentation:
    css: css/style.css
    logo: css/logo_en.png
    transition: 0
    self_contained: TRUE   
---

## Plan

A bit was pushed from the recap yesterday.

1. Functions
1. Loops and indexes
1. Apply family
1. `purrr` and `map`
1. Exercises
1. Split, apply, combine
1. Nested data and lists
1. Use case
1. Exercises again

## Why is it important?

- From [The tidy tools manifesto](https://cran.r-project.org/web/packages/tidyverse/vignettes/manifesto.html)

1. There are four basic principles to a tidy API:
1. Reuse existing data structures.
1. Compose simple functions with the pipe.
1. Embrace functional programming.
1. Design for humans.

# Functions

## Functions in general

- The backbone of R and all packages
- Make your own
- Complicated code & reusable? 
- Make it a function
- Remember: Good naming and pipe friendliness

## Existing functions

```{r}
min
survey::svydesign
```

## Existing functions

```{r}
`%in%`
```

## Make your own

1. Name function
1. Name input
1. Calculate output
1. Explicit return output

```{r, eval=FALSE}
my_function <- function(my_arguments){
  my_calculations
  return(my_return)
}
```

## Example

- Function that squares its input

```{r}
squared <- function(x){
  y <- x^2
  return(y)
}

squared(x = 1:3)
```

## Example - check input

```{r, error=TRUE}
squared("A")

squared_new <- function(x){
  stopifnot(is.numeric(x))
  y <- x^2
  return(y)
}
```

## Example - check input

```{r, error=TRUE}
squared_new("A")
squared_new(1:3)
```

## Example - several inputs

```{r}
new_function <- function(x,y,z){
  out <- (x + y)/z
  return(out)
}

new_function(
  x = c(1,2,10),
  y = c(4,7,9),
  z = c(8,10,2)
)
```

# Loops and indexes

## Indexes

- Two different ways to index
- Imagine a long train:
- `[]`: Selects which cars you are intered in
- `[[]]`: Selects whats **inside** the chosen car

## Indexes - Example

```{r}
x <- list(a = 1, b = "F", c = c(1,2))
x
```

## Indexes - Example

- Want the two first elements

```{r, eval=FALSE}
x[1:2] # By element number
x[c(TRUE,TRUE,FALSE)] # Logical for each element
x[-3] # By excluding unwanted element
```
```{r, echo=FALSE}
x[1:2]
```

- Names are kept
- Return is still a `list`

## Indexes - Example

- I only want the 3rd element

```{r}
x[[3]]
```

- No longer a `list`!

## Loops

- Repeating an action
- Manipulating list based on index
- Different loops:
- `for`: Loop over a sequence
- `while` Loop until a condition is no longer `TRUE` (be careful!)

## Loops - Example

```{r}
for (i in 1:2){
  print(i)
}

for (i in seq_len(3)) {
  print(i)
}
```

## Loops - Example

```{r}
x <- 1
while (x < 5){
  print(x)
  x <- x + 1
}
```

## Loops - Example

```{r, eval=FALSE}
x <- 1
while (x > 0){
  print(x)
  x <- x + 1
}
```

- Why is this a bad idea?

# Apply family

## Recap on apply/lapply

- Alternative to loops
- More functional programming
- `lapply`: When you want to evaluate the same function on a list of elements individually
- `apply`: When you want to calculate a function across a `matrix`

```{r, eval=FALSE}
lapply(X, FUN)
apply(X, MARGIN, FUN, ...)
```

## Example with lappy

- Calculate the sum of several `vectors` in a `list`:

```{r}
x <- list(
  a = 1:3,
  b = 10^(1:5)
)
```

## Example with lappy

```{r}
x
lapply(X = x, FUN = mean)
```

## Example with apply

```{r}
x <- matrix(data = (1:9)^2, nrow = 3)
x
```

## Example with apply

- Find the mean across the two dimensions:

```{r}
apply(X = x, MARGIN = 1, FUN = mean)
apply(X = x, MARGIN = 2, FUN = mean)
```

# Purrr and map

## The purrr package

- Another `tidyverse` package
- Alternative to `lapply`/`apply`
- Pipe-friendly
- Consistent output
- [Apply Functions Cheat Sheet](https://github.com/rstudio/cheatsheets/raw/master/purrr.pdf)

## Most important functions

- `map`: Very similar to lapply
- `map2`: Two argument version of `map`
- `pmap`: Unlimited number of arguments
- `walk`: Only for side effects
- `reduce`: Combining intermediate results
- `accumulate`: Combine but also keep intermediate results

## Control the output

- All `map` functions returns a list per default
- Use suffix to change that
- The list is then coerced into the desired class
- `_dbl`: Double or numeric
- `_int`: Integer
- `_chr`: Character
- `_lgl`: Logical
- `_dfr`: Data.frame (row binding)
- `_dfc`: Data.frame (column binding)

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```

## map function

```{r, eval=FALSE}
map(.x, .f, ...)
```

```{r}
x <- list(
  a = 1:3,
  b = 10^(1:5)
)
```

## map function

```{r}
map(.x = x, .f = mean)
x %>% map_dbl(.f = mean)
```

## map2 function

```{r}
x <- list(
  a = 1:3,
  b = 10^(1:5)
)

y <- list(
  a = 10:30,
  b = 7^(2:7)
)

f <- function(x1,x2){
  out <- sum(x2)/sum(x1)
  return(out)
}

```

## map2 function

```{r}
map2(.x = x, .y = y, .f = f)
```

## walk function

```{r, eval=FALSE}
map(.x = 1:2, .f = print)
walk(.x = 1:2, .f = print)
```

- What result do you expect?

## walk function

```{r}
map(.x = 1:2, .f = print)
walk(.x = 1:2, .f = print)
```

## walk function

- `print` function is a "side-effect" function that prints the object
- But invisible it returns the same object

```{r}
2 %>% 
  print() %>% 
  paste0()
```

- When you dont want the original output or to keep a list structure use `walk`
- Good when using plotting functions

## reduce and accumulate

- Result of two first are passed on as the first argument in the next evaluation
- List of `data.frames`

```{r}
df1 <- data.frame(id = 1:3)
df2 <- data.frame(id = 2:3)
df3 <- data.frame(id = 3)

l <- list(df1,df2,df3)
```

## reduce and accumulate

```{r}
l
```

## reduce and accumulate

- Join them all together with `inner_join`
- What do we expect?

## reduce and accumulate

- Join them all together with `inner_join`
- What do we expect?

```{r}
reduce(.x = l, .f = inner_join, by = "id")
```

## reduce and accumulate

- What about `accumulate?`

## reduce and accumulate

- What about `accumulate?`

```{r}
accumulate(.x = l, .f = inner_join, by = "id")
```

# Exercises

# Lambda functions

## Lambda functions in purrr

- `map` etc. requires a function
- Not all calculations need is relevant again
- Why explicitly assign a function to it?

## Tradtional way

```{r}
squared
map_dbl(.x = 1:3, .f = squared)
```

## Lambda way

```{r}
map_dbl(.x = 1:3, ~ .x^2)
```

- Formel interface/tilde expansion
- Easyish to read
- Recomdable instead of defining one-time use functions

## And it scales!

```{r}
map2_dbl(.x = 1:3, .y = 11:13, .f = ~ .x/.y)
new_function
pmap_dbl(.l = list(1:3, 11:13, 21:23), 
         .f = ~ (..1+..2)/..3)
```

# Split, apply, combine

## Strategy for data analysis

- **Split** unrelated data into chuncks
- **Apply** functions and estimations on each chuck
- **Combine** results for final analysis

## When is it good?

- Clustered data
- Better overview
- Larger and Big data
- Efficiency gain
- Improve structure in your analysis

# Nested data

## What is it?

- One way to implement split, apply, combine
- `tidyr` package
- [Apply Functions Cheat Sheet](https://github.com/rstudio/cheatsheets/raw/master/purrr.pdf)
- `lists` as columns in `tibbles`

## Example

- Nest the `mtcars` data based on cylinders

```{r}
df <- mtcars %>% 
  rownames_to_column("car") %>% 
  as_tibble()

df
```

## Example

- Easy to get an overview

```{r}
df_nested <- df %>% 
  nest(data = c(-cyl))

df_nested
```

## Example

- We can pluck one out

```{r}
df_nested$data[[1]]
```

## Example

- Manipulating a nested `tibble`

```{r}
df_nested_new <- df_nested %>% 
  mutate(hp_mean = map_dbl(.x = data, .f = ~ mean(.x$hp)))

df_nested_new
```

## Example

- Unnest it again

```{r}
df_nested_new %>% 
  unnest(cols = "data")
```

# Use case

## Sampling

Why can this be used?

- Simple sample
- Population: Persons
- Strata: Gender and Education
- Population size: 1,000,000
- Sample size: 5,000
- Proportional allocation

```{r}
N = 1e6
n = 5e3
```


Very simple version of Signe's problem.

## Make a population

```{r, message=FALSE}
library(tidyverse)
pop_0 <- tibble(
  ID = seq_len(N),
  GENDER = sample(x = c("F","M"), size = N, replace = TRUE),
  EDUCATION = sample(x = c("LOW","MEDIUM","HIGH"), 
                     size = N, replace = TRUE, prob = c(0.2, 0.5, 0.3))
)
```

## The new population

```{r}
pop_0
```

## Design

1. Split by strata
1. Calculate number in each
1. Calculate proportion of total
1. Calculate numbers to draw from each strata
1. Draw the sample from each
1. Combine them into the final sample

## Split by strata

- We use nesting!

```{r, message=FALSE}
pop_1 <- pop_0 %>% 
  nest(POP = c(-GENDER,-EDUCATION)) %>% 
  print()
```

## Calculate number in each

```{r}
pop_2 <- pop_1 %>% 
  mutate(N = map_int(.x = POP, .f = nrow)) %>% 
  print()
```

## Calculate proportion of total

```{r}
pop_3 <- pop_2 %>% 
  mutate(P = N/sum(N)) %>% 
  print()
  
```

## Calculate numbers to draw from each strata

```{r}
pop_4 <- pop_3 %>% 
  mutate(n = round(P * sum(N))) %>% 
  print()
```

## Draw the sample from each

```{r}
pop_5 <- pop_4 %>% 
  mutate(SAMPLE = map2(.x = POP, .y = n, 
                       .f = ~ sample_n(tbl = .x, size = .y))) %>% 
  print()
```

## Combine them into the final sample

```{r}
final_sample <- pop_5 %>% 
  select(GENDER,EDUCATION,SAMPLE) %>% 
  unnest(cols = SAMPLE)

nrow(final_sample)
```

## Final sample

```{r}
final_sample
```

## Could have been done in one step

```{r}
pop_final <- pop_0 %>% 
  nest(POP = c(-GENDER,-EDUCATION)) %>% 
  mutate(
    N = map_int(.x = POP, .f = nrow),
    P = N/sum(N),
    n = round(P * sum(N)),
    SAMPLE = map2(.x = POP, .y = n, 
                       .f = ~ sample_n(tbl = .x, size = .y))
    )
  
```

# Exercises again

# Done