---
title: "Functional Programming"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output:
  ioslides_presentation:
    css: css/style.css
    logo: css/logo_en.png
    transition: 0
    self_contained: TRUE   
---

## Plan

A bit was pushed from the recap yesterday.

1. Functions
1. Loops and indexes
1. Apply family
1. `purrr` and `map`
1. Exercises
1. Split, apply, combine
1. Nested data and lists
1. Use case
1. Exercises again

# Functions

## Functions in general

- The backbone of R and all packages
- Make your own
- Complicated code & reusable? 
- Make it a function
- Remember: Good naming and pipe friendliness

## Existing functions

```{r}
min
survey::svydesign
`%in%`
```

## Make your own

1. Name function
1. Name input
1. Calculate output
1. Explicit return output

```{r, eval=FALSE}
my_function <- function(my_arguments){
  my_calculations
  return(my_return)
}
```

## Example

- Function that squares its input

```{r}
squared <- function(x){
  y <- x^2
  return(y)
}

squared(x = 1:3)
```

## Example - check input

```{r, error=TRUE}
squared("A")

squared_new <- function(x){
  stopifnot(is.numeric(x))
  y <- x^2
  return(y)
}
```

## Example - check input

```{r, error=TRUE}
squared_new("A")
squared_new(1:3)
```

## Example - several inputs

```{r}
new_function <- function(x,y,z){
  out <- (x + y)/z
  return(out)
}

new_function(
  x = c(1,2,10),
  y = c(4,7,9),
  z = c(8,10,2)
)
```

# Loops and indexes

## Loops

## Indexes

- Two different ways to index
- Imagine a long train:
- `[]`: Selects which cars you are intered in
- `[[]]`: Selects whats **inside** the chosen car

## Indexes - Example

```{r}
x <- list(a = 1, b = "F", c = c(1,2))
x
```

## Indexes - Example

- Want the two first elements

```{r, eval=FALSE}
x[1:2] # By element number
x[c(TRUE,TRUE,FALSE)] # Logical for each element
x[-3] # By excluding unwanted element
```
```{r, echo=FALSE}
x[1:2]
```

- Names are kept
- Return is still a `list`

## Indexes - Example

- I only want the 3rd element

```{r}
x[[3]]
```

- No longer a `list`!

# Purrr and map

# Exercises

# Split, apply, combine

# Nested data and lists

# Use case

# Exercises again

# Done