---
title: "Functional Programming - Exercises 1"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output: html_document
---

None of the problems depend on each other, so if you are stuck just continue to the next.

If in need of a hint just ask or consult one of the many cheat sheets ;)

Each problem can be solved in different ways.

## Problem 1

Create your own function to calculate the cummulative sum of a vector.
For inspiration the base R function `cumsum` works like this:

```{r}
x <- 1:10
x
cumsum(x)
```

The function should return the same.

**Solution 1**

```{r}
x <- 1:10
y <- x
for (i in 2:10) {
  y[[i]] <- x[[i]] + y[[i-1]]
}
x
y
cumsum(x)

cumulative_sum <- function(x){
  y <- x
  for (i in 2:length(x)) {
    y[[i]] <- x[[i]] + y[[i-1]]
  }
  return(y)
}

cumulative_sum(1:7)
```

**Solution 2**

```{r}
library(purrr)
accumulate(.x = 1:7, .f = sum)

cumulative_sum <- function(x){
  y <- accumulate(.x = x, .f = sum)
  return(y)
}

cumulative_sum(1:9)
```

## Problem 2

Make a function that given two inputs of numeric vectors returns the pairwise mean.
E.g. the following input:

```{r}
(x <- c(1,5,9))
(y <- c(2,4,0))
```

Should return:

```{r}
c(1.5,4.5,4.5)
```

**Solution 1**

```{r,error=TRUE}

pmean <- function(x,y){
 stopifnot(length(x) == length(y))
  stopifnot(is.numeric(x) & is.numeric(y))
  out <- numeric(length = length(x))
  for (i in seq_len(length(x))) {
    out[[i]] <- mean(x[[i]],y[[i]])
  }

return(out) 
}

pmean(x,y)
pmean(x,"a")
pmean("A",1)
```

**Solution 2**

```{r}
x
y
map2(.x = x, .y = y, .f = c) %>% 
  map_dbl(.f = mean)
```

## Problem 3

I have uploaded 5 different csv files with data on cars to the webpage (cars_1.csv - cars_5.csv).
Download and read them all into R using functional programming, and combine them into one common `data.frame`.

**Solution 1**

## Problem 4

Create a histogram of each numeric variable in the cars data.
It does not mather if you use `mtcars` or the data from Problem 3, nor wether the histogram is produced with base `hist` or produced with `ggplot`.

**Solution 1**

## Problem 5

Calculate the minimum for each column in a `data.frame`. Either though a small script or a function.

For example the data:

```{r, message=FALSE}
library(tidyverse)
tibble(x = 1:3,
       y = c(500,60,70))
```

Should give the result:

```{r}
tibble(x = 1, y = 60)
```

**Solution 1**

## Problem 6

Can you expand the idea in problem 2 into creating a function or small code chunck, that calculates the "paiwise" (or index-wise) mean across any given number of vectors?
Inspired by the `pmin` function:

```{r}
pmin(
  c(1,3,10,5), 
  c(4,2,11,3))

pmin(
  c(1,3,10,5),
  c(4,2,11,3),
  c(7,-2,9,8))
```

**Solution 1**

