---
title: "Functional Programming - Exercises 2"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output: html_document
---

There is four problems 

## Problem 7

On the `mtcars` estimate the following linear model:

$qsec = a \cdot hp + b \cdot wt$

So where the time in seconds to complete a quarter-mile (`qsec`) is a linear function of horsepower (`hp`) and weight in 1,000 lbs (`wt`). Hint: use the `lm` function.

Afterwards extract the coefficient (`a` & `b` in the equation) into seperate variables.
Hint: Experiment with the `coef` function.

**Solution**

Estimate model

```{r,message=FALSE}
library(tidyverse)

df <- as_tibble(mtcars)
df

model <- lm(qsec ~ hp + wt, data = df)

summary(model)
```

Extract coefficient

```{r}
coef(model)
a <- coef(model)[[2]]
b <- coef(model)[[3]]
a
b
```

## Problem 8

Divide the `mtcars` data in subsets based on the number of gears (`gear`) each car has.

Now estimate the same model from Problem 7, but this time estimate one individual model for each of these subset.

Finally compare the results.

**Solution**

```{r}
df <- mtcars %>% 
  as_tibble() %>% 
  nest(data = c(-gear)) %>% 
  mutate(
    MODEL = map(.x = data, .f = ~ lm(qsec ~ hp + wt, data = .x)),
    COEF = map(.x = MODEL, .f = coef),
    a = map_dbl(.x = COEF, .f = ~ .x[[2]]),
    b = map_dbl(.x = COEF, .f = ~ .x[[3]])
  )

df

summary(df$MODEL[[1]])
df$COEF
```

## Problem 9

Using the `mydata_example` data from yesterday:

```{r, message=FALSE}
library(tidyverse)
load("mydata_example.Rda")

mydata_example <- as_tibble(mydata_example)
mydata_example
```

Draw a simple stratified sample with the following specifications:

- Sample size: 200
- Strata: `region` & `sector`
- Proportional allocation

**Solution**

```{r}
pop_final <- mydata_example %>% 
  nest(POP = c(-region,-sector)) %>% 
  mutate(
    N = map_int(.x = POP, .f = nrow),
    P = N/sum(N),
    n = round(P * 200),
    SAMPLE = map2(.x = POP, .y = n, 
                       .f = ~ sample_n(tbl = .x, size = .y))
    )

pop_final
sum(pop_final$n)

pop_final %>% 
  select(region,sector,SAMPLE) %>% 
  unnest(cols = SAMPLE)
```

## Problem 10

In stead of proportional allocation can you draw a sample with optimal allocation based on the variance of `turnover_2019`?
Otherwise the same specifications as Problem 9.

**Solution**

```{r}
pop_final <- mydata_example %>% 
  nest(POP = c(-region,-sector)) %>% 
  mutate(
    N = map_int(.x = POP, .f = nrow),
    SD = map_dbl(.x = POP, .f = ~ sd(.x[["turnover_2019"]])),
    P = N * SD / sum(SD * N),
    n = round(P * 200),
    SAMPLE = map2(.x = POP, .y = n, 
                       .f = ~ sample_n(tbl = .x, size = .y))
    )

sum(pop_final$n)

pop_final
```

