---
title: "Functional Programming - Exercises 1"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output: html_document
---

None of the problems depend on each other, so if you are stuck just continue to the next.

If in need of a hint just ask or consult one of the many cheat sheets ;)

Each problem can be solved in different ways.

## Problem 1

Create your own function to calculate the cummulative sum of a vector.
For inspiration the base R function `cumsum` works like this:

```{r}
x <- 1:10
x
cumsum(x)
```

The function should return the same.

**Solution 1**

## Problem 2

Make a function that given two inputs of numeric vectors returns the pairwise mean.
E.g. the following input:

```{r}
(x <- c(1,5,9))
(y <- c(2,4,0))
```

Should return:

```{r}
c(1.5,4.5,4.5)
```

**Solution 1**

## Problem 3

I have uploaded 5 different csv files with data on cars to the webpage (cars_1.csv - cars_5.csv).
Download and read them all into R using functional programming, and combine them into one common `data.frame`.

**Solution 1**

## Problem 4

Create a histogram of each numeric variable in the cars data.
It does not mather if you use `mtcars` or the data from Problem 3, nor wether the histogram is produced with base `hist` or produced with `ggplot`.

**Solution 1**

## Problem 5

Calculate the minimum for each column in a `data.frame`. Either though a small script or a function.

For example the data:

```{r, message=FALSE}
library(tidyverse)
tibble(x = 1:3,
       y = c(500,60,70))
```

Should give the result:

```{r}
tibble(x = 1, y = 60)
```

**Solution 1**

## Problem 6

Can you expand the idea in problem 2 into creating a function or small code chunck, that calculates the "paiwise" (or index-wise) mean across any given number of vectors?
Inspired by the `pmin` function:

```{r}
pmin(
  c(1,3,10,5), 
  c(4,2,11,3))

pmin(
  c(1,3,10,5),
  c(4,2,11,3),
  c(7,-2,9,8))
```

**Solution 1**

