---
title: "Introduction & Repetition"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output:
  ioslides_presentation:
    css: css/style.css
    logo: css/logo_en.png
    transition: 0
    self_contained: TRUE   
---

# Welcome

## Revised programme {.smaller}

| **Day**   |  **Time**     | **Purpose/Details** | **Material** | **Lecturer** | 
|-----------|---------------|---------------------|--------------|--------------|
| Monday    | 15:00 - 16:00 | Repetition                          | A | Aksel   |
| Tuesday   | 9:00 - 12:00  | Sampling                            | B | Signe   |
|           | 13:00 - 15:30 | Estimation and calibration          | C | Signe   |
| Wednesday | 9:00 - 12:00  | Functional programmering            | D | Aksel   |
|           | 13:00 - 15:30 | Prepare and analyse large data      | E | Aksel   |
| Thursday  | 9:00 -  XX:XX | Further work with large data        | F | Aksel   |

## Format
- Same as the last courses
- Lectures followed by practicals
- Materials are on the homepage
  - "Impracticals" are the versions without solutions, "Practicals" the ones with
- Please do ask questions!

## Repitition

1. A few base R essentials
1. R Objects
1. Tidyverse
1. Importing data into R

# A few base R essentials

## Native classes

- Types:
  - `vector`
  - `matrix`
  - `list`
  - `data.frame`
- Classes:
  - `numeric`
  - `integer`
  - `character`
  - `Date`
  - `POSIXct` or datetime
  
## Functions for tomorrow

- `aggregate`: Summarise data
- `cut`: Divide data

## Aggregate

- Summarise a data frame
- Formula based

```{r}
df <- data.frame(group = c(1,1,1,2,2), x = 1:5)
df
```

## Aggregate

- Summarise a data frame
- Formula based

```{r}
aggregate(x ~ group, data = df, FUN = sum)
```

## Cut

- Divides the range
- Recodes numeric to factor

```{r}

x = runif(n = 10, min = 0, max = 1)
x

cut(x = x, breaks = c(0,0.25,0.75,1))
```

# R objects

# Into the Tidyverse

## What is it?

- From Rstudio: https://www.tidyverse.org/
- Common (and opinionated) collection of packages
- Pipes
- `dplyr`
- `tibbles`
- `ggplot2`

# Importing data into R

## Very common task

- Native object class for data is the `data.frame`
- But data is always from somewhere else
- Most commonly: CSV & Excel files
- But also supprt for other formats: SAS, SPSS and many more

## Importing CSV

```{r,  include=FALSE}

# How the data was made

library(tidyverse)
x <- mtcars %>% 
  rownames_to_column("car") %>% 
  head(5)

write.csv(x = x, file = "mtcars5.csv", row.names = FALSE)
openxlsx::write.xlsx(x = x, file = "mtcars5.xlsx")
```


```{r}
df <- read.csv(file = "mtcars5.csv")
df
```

## Check the output

- Is it to be expected?
- Check the column classes
- Very rarely you actually want factors!

```{r}
as_tibble(df)
```

## Two options

1. Control with arguments to the base R function
1. Use the excellent ´readr´ package (in general a really good idea!)

## Option 1

```{r}
as_tibble(read.csv("mtcars5.csv", stringsAsFactors = FALSE))
```

## Option 2

```{r}
df <- readr::read_csv("mtcars5.csv")
```

## Option 2

```{r}
df
```

## For Excel

- Excellent `openxlsx` package
- Both to write and read

```{r}
df <- openxlsx::read.xlsx(xlsxFile = "mtcars5.xlsx")
as_tibble(df)
```

## For everything else

- The `haven` package has all the foreign formats:

```{r, echo=FALSE, results='asis'}
getNamespaceExports("haven") %>% 
  str_subset("read_") %>% 
  paste0("- ",.,"\n") %>% 
  paste(collapse = ' ') %>% 
  cat()
```

