---
title: "Analysing large data"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output:
  ioslides_presentation:
    css: css/style.css
    logo: css/logo_en.png
    transition: 0
    self_contained: TRUE   
---

## Plan

- Quick intro to analysing "large" data in R
- Apply tools from yesterday
- The `disk.frame` package
- Enables manipulation of "large" data
- Use case with AIS data
- Exercises & walkthrough

# Large data

## What is it?

- *Normal data*: Easily fit in RAM - everyday R data
- *Large data*: Bigger than RAM - but fits on disk
- *Big data*: Too big for disk - need special databases

## Special problems

- Does not fit in RAM
- Can´t analyse data as normal
- Split, apply, combine is the solution!

## Possible tools

- Functional programming from yesterday
   - `map`
   - `reduce`
   - `...`
- Special packages
   - `disk.frame`
   
# AIS data

## What is AIS?

- **A**utomatic **I**dentification **S**ystem
- Positional data on ships in Denmark
- An observation per 10 second per ship
- Our variables of interest:
   - `SOG`: Speed over ground
   - `Longitude`
   - `Latitude`
- Data for December 1-5 2019
- Seperate CSV files

## The data

```{r, message=FALSE}
library(tidyverse)
tibble(FILE = list.files(pattern = "aisdk")) %>% 
  arrange(FILE) %>% 
  mutate(SIZE_MB = map_dbl(.x = FILE, file.size)/1024^2)
```

## Read one in

```{r}
df <- readr::read_csv(file = "aisdk_201902.csv")
```

## Inspect data

```{r}
nrow(df)
ncol(df)
```

## Inspect data

```{r}
names(df)
```

## Inspect data

```{r}
as_tibble(df)
```

## Only relevant

```{r}
df_relevant <- df %>% 
  as_tibble() %>%
  select(TIMESTAMP = `# Timestamp`, MMSI, SOG, Latitude, Longitude)

df_relevant
```

## One ship has many observations

```{r}
df_relevant %>% filter(MMSI == 2194005)
```

## Summary statistics

```{r, message=FALSE}
pryr::object_size(df)
pryr::object_size(df_relevant)
summary(df_relevant)
```

## Learnings

- One day is an okay size
- Only keep relevant information
- If we had all days we might be in trouble memory wise

# disk.frame

## The `disk.frame` package

- [diskframe.com](https://diskframe.com/index.html)
- Purpose: To manipulate data larger than RAM
- The idea:
   1. Split up a larger-than-RAM dataset into chunks
   1. Provide a convenient API to manipulate these chunks
- Split & apply!

## How it works

- Splits data into small chunks saved on the disk
- Uses parrallel processes to manipulate them fast!

## How to learn it

- It is relatively new
- Very good webpage
- With very good examples and vignettes
- Go through them next time you need `disk.frame`

## Setup

```{r}
library(disk.frame)
```

## Setup

```{r}
# this will set up disk.frame with multiple workers
setup_disk.frame()
# this will allow unlimited amount of data to be passed from worker to worker
options(future.globals.maxSize = Inf)
```

## Create disk.frame

```{r}
df_disk <- as.disk.frame(df_relevant)

df_disk
```

## Directly from CSV

From one CSV:

```{r, eval=FALSE}
df_disk_csv <- csv_to_disk.frame(infile = "aisdk_201902.csv")
```

From several in one go:

```{r, eval=FALSE}
df_disk_csv <- csv_to_disk.frame(
  infile = c("aisdk_201901.csv","aisdk_201902.csv")
  )
```

## How to manipulate

- Most `dplyr` verbs are defined
- Check the webpage
- Manipulate the chunks yourself and combine afterwars
- Functional programming again!

## New functions

- `delayed(.x , .f, ...)`
- `chunk_mutate`
- `chunk_group_by`
- `chunk_summarise`
- `collect`
- `collect_list`

## Examples

Calculate the number of observations in each chunk

```{r}
df_disk %>% 
  delayed(.f = nrow)
```

## Examples

```{r}
df_disk %>% 
  delayed(.f = nrow) %>% 
  collect_list()
```

## Examples

Calculate mean speed of all observations

```{r}
df_disk %>% 
  chunk_summarise(SOG_MEAN = mean(SOG, na.rm = TRUE)) %>% 
  collect()
```

## Examples

Calculate mean speed of all observations

```{r}
df_disk %>% 
  chunk_summarise(N = n(),
                  SOG_MEAN = mean(SOG, na.rm = TRUE)) %>% 
  collect() %>% 
  print() %>% 
  summarise(SOG_MEAN_TOTAL = sum(N * SOG_MEAN)/sum(N) )
```

# Exercise
