---
title: "Analysing large data - Solution"
author: "Aksel Thomsen"
date: "Statistical Programming with R"
output: html_document
---

In this exercise we will try to recreate a map of sailing rutes in Denmark.
It is a good example on how you can gain a lot of information by applying a quite simple analysis to a large amount of data.

## Problem 1

Here we want to only keep the relevant information in the data.
The analysis requires the position of the moving ships.

One way of restricting the data is as follows:

1. Read any one of the csv files into R
1. Select only the relevant variables
1. Filter so you only have the relevant observations

**Solution**

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
df1 <- readr::read_csv("aisdk_201902.csv")
```

```{r}
head(df1)
```

```{r}
df2 <- df1 %>% 
  select(Latitude,Longitude,SOG) %>% 
  filter(!is.na(SOG),SOG>0)
```

```{r}
head(df2)
```

## Problem 2

Can you apply the same restrictions to all five CSV files and combine them to one data ready for analysis?

If you are running out of RAM use the `n_max` argument to `readr::read_csv` to limit how many observations you include. For this specific problem it is okay not include all.

**Solution**

```{r}
ais_files <- list.files(pattern = "aisdk")

ais_files
```

I use `n_max=1000` so I don't run out of memory.

```{r, warning=FALSE, message=FALSE}
df1_all <- purrr::map(.x = ais_files, .f = readr::read_csv, n_max = 1000)
```

```{r}
map(df1_all, head, 5)
```

```{r}

df2_all <- df1_all %>% 
  purrr::map(
    ~ .x %>% 
      select(Latitude,Longitude,SOG) %>% 
      filter(!is.na(SOG),SOG>0)
  )

map(df2_all, head, 5)

df_combined <- bind_rows(df2_all)

head(df_combined)
```

## Problem 3

Create a `disk.frame` in stead based on the CSV files.
Remember to run the setup code after loading the package.

Hint: To save developing time start with only using one CSV file or just use a single one.

- How many chunks is it divided in?
- How many observations do they have?

**Solution**

```{r,message=FALSE}
library(disk.frame)
# this will set up disk.frame with multiple workers
setup_disk.frame()
# this will allow unlimited amount of data to be passed from worker to worker
options(future.globals.maxSize = Inf)
```

```{r}
ais_files <- list.files(pattern = "aisdk")

ais_df <- csv_to_disk.frame(infile = ais_files, outdir = "/ais.df")

ais_df
```

```{r}
n_obs <- ais_df %>% 
  delayed(.f = nrow)

collect_list(n_obs)
```

## Problem 4

Apply the same restrictions from problem 1 & 2 to the `disk.frame`.

**Solution**

```{r}
ais_df_restricted <- ais_df %>% 
  select(Latitude,Longitude,SOG) %>% 
  filter(!is.na(SOG),SOG>0)

ais_df_restricted

# Data is not collected until the end to save computing time and memory
# collect(ais_df_restricted)
```

## Problem 5

- Define a 1000 x 1000 meter grid
- A simple way is just to divide the coordinates with 1000 and round them
- Calculate the number of observations in each grid over the 5 days?

**Solution**

First i calculate the grids.
In stead of 1000 x 1000 meter grids, I calculate rounded to digits degrees.
Sorry for the confusion. The two grid sized are not the same, but the the technique is equilivant.

```{r}
head(ais_df_restricted)

ais_df_grid <- ais_df_restricted %>% 
  mutate(X = round(Longitude, digits = 2),
         Y = round(Latitude, digits = 2))

head(ais_df_grid)
```

Summarise and collect the results for each chunk.

```{r}
ais_df_sum <- ais_df_grid %>% 
  chunk_group_by(X,Y) %>% 
  chunk_summarise(N = n())

ais_df_sum <- collect(ais_df_sum)
```

Summarise the chunk results.

```{r}
ais_df_sum

ais_df_sum2 <- ais_df_sum %>% 
  group_by(X,Y) %>% 
  summarise(N = sum(N)) %>% 
  ungroup()

ais_df_sum2

ais_df_sum2 %>% 
  arrange(desc(N)) %>% 
  head(10)
```

## Problem 6

We can plot Denmark as follows (from the first course):

```{r, message=FALSE}
library(tidyverse)
library(sf)

denmark <- st_read("DK_map.shp")

ggplot() +
  geom_sf(data = denmark) +
  theme_minimal()
```

Use the data from problem 5 to make a heatmap of the most transited waters.
Hint: Plot the grids on top colour encoded.

**Solution**

First I filter the grid data removing the points outside Denmark.

```{r}
ais_df_filtered <- ais_df_sum2 %>% 
  filter(X > 7, X < 16, Y > 54, Y < 58)
```

Simple plot:

```{r}
ggplot(data = ais_df_filtered) +
  geom_point(mapping = aes(x = X, y = Y, colour = log(N))) +
  scale_colour_viridis_c()
```

Convert to `sf` object, making it easier to overlap with the map of Denmark. 

```{r}
ais_sf <- ais_df_filtered %>% 
  st_as_sf(coords = c("X","Y")) %>% 
  st_set_crs(value = 4326) #The reference of the coordinate system
```

```{r}
ais_sf
```

How does it look now?

```{r}
ggplot() +
  geom_sf(data = ais_sf, mapping = aes(colour = N)) +
  scale_colour_viridis_c()
```

First attempt at combining them:

```{r}
ggplot() +
  geom_sf(data = denmark) +
  geom_sf(data = ais_sf, mapping = aes(colour = log(N))) +
  scale_colour_viridis_c() +
  theme_minimal()
```

There is a lot of points inside the country. A simple way of making it look nicer is to change the order of plotting, so that the map of Denmark is plotted last in stead:

```{r}
ggplot() +
  geom_sf(data = ais_sf, mapping = aes(colour = log(N))) +
  geom_sf(data = denmark, fill = "grey") +
  scale_colour_viridis_c() +
  theme_minimal()
```

Note: In order to make a proper grid plot we would need to plot squared polygons in stead of simple dots, but the interpretation and method would be the same.